
#  Prerequisites

The general tools you'll need in the course:
* R, i.e. the program
* Rstudio, i.e. the nice graphical interface that makes things easier
* tidyverse: the collection of packages that make the data cleaning, wrangling and analysis smoother
* ggplot2: visualization package by the tidyverse main developer
* BioConductor, i.e. the place where most of the bioinformatics tools live


It's easier during the course if you have a general understanding of the inner works of R, tidyverse and ggplot2. There are great openly available materials that introduce the topic, and I highly recommend watching at least these:

1. Videos 2-5, 8-14 from [Marin Stats Lectures](https://www.youtube.com/playlist?list=PLqzoL9-eJTNARFXxgwbqGo56NtbJnB37A)
* While you are watching the videos:
+ Install R and RStudio
+ Download an example data that I'll send you
+ Calculate a correlation between two numeric variables

You can use other materials in the internet, and effective use of search engines is the key for problem solving. For example, [STHDA](http://www.sthda.com) and [Stack Overflow](https://stackoverflow.com) are valuable resources.

2. Understand the basic anatomy of the general data science tools: [tidyverse](https://www.youtube.com/watch?v=nRtp7wSEtJA) (starting from 1:33) and [ggplot2](https://www.youtube.com/watch?v=FdVy57oGJuc) (starting from 0:52)
* Visualize the relationship between two numeric variables
* Generate similar plots separately for males and females

3. Install the specialist tools:
* [Bioconductor](https://www.bioconductor.org/install/), [video instruction](https://www.youtube.com/watch?v=18FyH5pKkuM)
* The stable version of [mia package](https://microbiome.github.io/OMA/packages.html)

**Do not hesitate to ask for support.**

> Congrats, now you have successfully laid the base for any data analysis projects!


**If you want to read more** I recommend the following:
* (Data science basics in R](https://r4ds.had.co.nz)
* (Modern Statistics for Modern Biology)[https://www.huber.embl.de/msmb/] open access book by Susan P. Holmes and Wolfgang Huber


## Support and resources

For additional reading and online material, see [Material](material.html) section

For online support on installation and other matters, you can join us at:

 * Users: [miaverse Gitter channel](https://gitter.im/microbiome/miaverse?utm_source=badge&utm_medium=badge&utm_campaign=pr-badge&utm_content=badge)
 * Local: Slack channel



## Installing and loading the required R packages

This section shows how to install and load all required packages into
the R session. Only uninstalled packages are installed.

```{r warning = FALSE, message = FALSE}
# List of packages that we need from cran and bioc 
cran_pkg <- c("BiocManager", "bookdown", "dplyr", "ecodist", "ggplot2", 
              "gridExtra", "kableExtra", "knitr", "scales", "vegan")
bioc_pkg <- c("ANCOMBC", "ape", "DESeq2",  "DirichletMultinomial", "mia", "miaViz")

# Gets those packages that are already installed
cran_pkg_already_installed <- cran_pkg[ cran_pkg %in% installed.packages() ]
bioc_pkg_already_installed <- bioc_pkg[ bioc_pkg %in% installed.packages() ]

# Gets those packages that need to be installed
cran_pkg_to_be_installed <- setdiff(cran_pkg, cran_pkg_already_installed)
bioc_pkg_to_be_installed <- setdiff(bioc_pkg, bioc_pkg_already_installed)
```

```{r warning = FALSE, message = FALSE}
# If there are packages that need to be installed, installs them from CRAN
if( length(cran_pkg_to_be_installed) ) {
   install.packages(cran_pkg_to_be_installed)
}
```

```{r warning = FALSE, message = FALSE}
# If there are packages that need to be installed, installs them from Bioconductor
if( length(bioc_pkg_to_be_installed) ) {
   BiocManager::install(bioc_pkg_to_be_installed, ask = F)
}
```
 
Now all required packages are installed, so let's load them into the session.
Some function names occur in multiple packages. That is why miaverse's packages
mia and miaViz are prioritized. Packages that are loaded first have higher priority.

```{r warning = FALSE, message = FALSE}
# Reorders bioc packages, so that mia and miaViz are first
bioc_pkg <- c(bioc_pkg[ bioc_pkg %in% c("mia", "miaViz") ], 
              bioc_pkg[ !bioc_pkg %in% c("mia", "miaViz") ] ) 

# Loading all packages into session. Returns true if package was successfully loaded.
loaded <- sapply(c(bioc_pkg, cran_pkg), require, character.only = TRUE)
as.data.frame(loaded)
```
